ESRC SDAI Application

### 2019-06-21 case_for_support.tex
Anna update

 \textcolor{red}{\textit{Short intro about aims, highlighting added value.}}
In our first research project, we will be using the CSEW and WMP datasets to investigate three main research questions. First, we will use the CSEW and WMP to identify the demographic and socio-economic characteristics of the victims of domestic abuse, and the WMP to characterise perpetrators of reported cases. Exploring the characteristics of groups vulnerable to domestic abuse, and the true extent of the problem of underreporting is crucial for designing effective interventions and prevent potential cases of domestic homicide. Second, we will estimate the true extent of domestic abuse victimisation. Given the multifaceted nature of domestic abuse and the problem of underreporting, it is very hard to reliably estimate the extent of the problem and the societal costs of domestic abuse. We wish to do so by combining the strengths of the two datasets (that is, the CSEW provides a more reliable measure of the true rate of victimisation across different offence types, while the WMP provides unique information on whether various types of offences were domestic abuse related). Finally, we will estimate the extent of police mis-recording domestic abuse. Improving the police's response to reported incidents is key in encouraging victims to report domestic abuse. By contrasting estimated levels from reported incidents within the West Midlands from the two datasets, we can identify any reporting gaps between the two sources of data that might arise from police mis-recording.

@@

Aim 3. \href{https://www.justiceinspectorates.gov.uk/hmicfrs/news/news-feed/little-evidence-of-improvements-in-crime-recording-at-west-midlands-police/}{HMICFRS}

@@

\textbf{Aim 1.} First, using the CSEW data, we will conduct a statistical analysis to provide a deeper understanding of the demographic and socio-economic factors and their interactions in predicting domestic abuse victimisation. We will first use a logistic regression approach to explore the factors predicting individual-level domestic abuse victimisation, and the type of abuse suffered (physical abuse, threats, sexual abuse). The CSEW provides a wealth of information on the characteristics of the respondent including age, sex, marital status, number of children, ethnicity, education, employment, income, benefit history, physical and mental health, frequency of going out, house and car ownership, well-being, drug and alcohol use, and we will evaluate the importance of each of these variables and their interaction in predicting our outcome variables. \textit{Could we also use a e.g. a random forest classification algorithm and look at variable importance?}  These analyses will represent the first extensive quantitative investigation of the individual-level predictors of domestic abuse victimisation in England, using a national-level, large representative sample like the CSEW. We are particularly interested in the association between experiences of domestic abuse and financial resilience, as previous research has suggested that access to resources is an important factor in escaping the cycle of abuse, especially for women with children.

Second, we can validate some of these findings using the WMP data. This dataset contains information about the sex and age of victims. We can contrast the sex-age victimisation numbers from the WMP with corresponding population estimates to validate the findings of the CSEW\textit{(we were initially talking about using victims of burglary, but maybe this is better?)}. We will also use the WMP data on the offender's sex and age to provide a characterisation of domestic abuse perpetrators.

@@

\textbf{Aim 3.} Police forces in England, including the West Midlands Police, have previously been criticised for mis-recording violent cases of violent crime, including \href{https://www.bbc.co.uk/news/uk-england-468676578}{domestic abuse.} Mis-recording these already heavily underreported crimes decrease victim's trust in the Criminal Justice System and their willingness to report subsequent incidents. By contrasting the extrapolated estimates from the CSEW (which is representative sample) with WMP data on the true number of reported cases, we can estimate the true extent of the mis-recording of domestic abuse cases by the WMP. Given the repeated cross-sectional nature of the CSEW, and the longitudinal nature of the WMP, we can quantify the proportion of mis-recorded incidents over time. We can complement this investigation with analysing the self-reported outcome of incidents that have been reported to the police from the CSEW.

@@

does the level of harm depend on the length of time spent in the relationship

@@

\textbf{Aim 1.} First, we will use the CSEW, and a multinomial logistic approach to understand how the level of harm suffered depends on the individual-level characteristics of the victim (especially focusing on the issue of separation and divorce, as it has proven to be a significant risk factor predicting serious harm) and the co-morbidity with other types of victimisation (other offence types). The CSEW asks whether the respondent needed to see a doctor, nurse or other health worker, or needed to take time off work because of the abuse.

@@

We can predict incidents with serious harm, using the type of offence occurred previously. number of days elapsed. This will allow us to identify and commonalities in the pattern of domestic abuse cases that lead to escalation in the form of serious harm. We can also identify and characterise high-risk perpetrators who breach court orders designed to protect the victims from further harms.

@@

\textbf{Aim 2.} First, we will use the CSEW, and a logistic approach to understand how the decision to report to the police (or seeking any external help) depends on the individual-level characteristics of the victim (especially focusing on mental and physical health and financial independence) and the abuse suffered.

>>>>

 \textcolor{red}{\textit{Short intro about aims, highlighting added value.}}
In our first research project, we will be investigating three main research questions. First, we will use the CSEW and WMP datasets to identify the demographic and socio-economic characteristics of victims of domestic abuse, and the WMP to characterise perpetrators of domestic abuse cases reported to the police. Second, we will estimate the true extent of domestic abuse victimisation. Given the multifaceted nature of domestic abuse, and the pervasive problem of underreporting, it is very hard to reliably estimate the extent of domestic abuse victimisation and the subsequent societal costs. We wish to do so by combining the strengths of the two datasets (that is, the CSEW provides a more reliable measure of the true rate of victimisation across different offence types, while the WMP provides unique information on whether various types of offences were domestic abuse-related). Exploring the risk factors in domestic abuse victimisation, and the true extent of the problem of underreporting is crucial for designing effective interventions to reduce the risk of escalation that may culminate in domestic homicide. Finally, we will estimate the extent of police mis-recording of domestic abuse. Increasing victim's trust in the police through improving law enforcement response to reported incidents is key in encouraging victims to report domestic abuse and prevent further harm. By contrasting estimated levels from reported incidents within the West Midlands from the two datasets, we can identify any reporting gaps between the two sources of data that might arise from police mis-recording.

@@

Aim 3. \href{https://www.justiceinspectorates.gov.uk/hmicfrs/news/news-feed/little-evidence-of-improvements-in-crime-recording-at-west-midlands-police/}{HMICFRS about West Midlands mis-recording}

@@

\textbf{Aim 1.} First, using the CSEW data, we will conduct a series of statistical analyses to provide a deeper understanding of the demographic and socio-economic factors and their interactions in predicting domestic abuse victimisation. Using a logistic regression approach, we will explore the factors predicting individual-level domestic abuse victimisation, depending on the type of abuse suffered (physical abuse, threats, sexual abuse). The CSEW provides a wealth of information on the characteristics of the respondent, including age, sex, marital status, number of children, ethnicity, education, employment, income, benefit history, physical and mental health, frequency of going out, house and car ownership, self-reported well-being, frequency of drug and alcohol use. Using this wide range of individual-level information, we will evaluate the importance of each of these variables and their interaction in predicting our outcome variables. \textit{Could we also use a e.g. a random forest classification algorithm and look at variable importance?}  These analyses will represent the first extensive quantitative investigation of the individual-level predictors of domestic abuse victimisation in England, using a national-level, large representative sample like the CSEW. We are particularly interested in the association between experiences of domestic abuse and financial resilience, as previous research has suggested that access to resources is an important factor in escaping the cycle of abuse, especially for women with children.

Second, we can validate some of these findings using the WMP data. This dataset contains information about the sex and age of victims. We can contrast the sex-age victimisation numbers from the WMP with corresponding population estimates to validate the findings of the CSEW\textit{(we were initially talking about using victims of burglary, but maybe this is better?)}. We can complement this analysis with data on the offender's sex  and age to estimate the prevalence of different forms of reported domestic abuse (child to parent violence, parent to child violence, intimate partner violence) and contrast it with estimates based on self-reported data from the CSEW.

@@

\textbf{Aim 3.} Police forces in England, including the West Midlands Police, have previously been criticised for mis-recording violent cases of violent crime, including \href{https://www.bbc.co.uk/news/uk-england-468676578}{domestic abuse.} Mis-recording these already heavily underreported crimes decrease victim's trust in the Criminal Justice System and their willingness to report subsequent incidents. By contrasting the extrapolated estimates from the CSEW (which is a representative sample) with WMP data on the true number of reported cases, we can estimate the true extent of the mis-recording of domestic abuse cases by the WMP. Given the repeated cross-sectional nature of the CSEW, and the longitudinal nature of the WMP, we can quantify the proportion of mis-recorded incidents over time. We can complement this investigation with analysing the self-reported outcome of incidents that have been reported to the police from the CSEW.

@@

does the level of harm depend on the length of time spent in the relationship, separation,

@@

\textbf{Aim 1.} First, we will use the CSEW, and a multinomial logistic approach to understand how the level of harm suffered (threats, minor or serious injury) depends on the individual-level characteristics of the victim (especially focusing on the issue of separation and divorce, as evidence suggests it is a significant risk factor predicting serious harm). We will also explore whether non domestic abuse-related victimisation is predictive of serious harm. We will know the level of harm suffered, because the CSEW asks whether the respondent needed to see a doctor, nurse or other health worker, or needed to take time off work because of the abuse.

@@

We can predict incidents with serious harm, using the type of offence occurred previously, criminal history of the offender. number of days elapsed. This will allow us to identify and commonalities in the pattern of domestic abuse cases that lead to escalation in the form of serious harm. We can also identify and characterise high-risk perpetrators who breach court orders designed to protect the victims from further harms.

@@

\textbf{Aim 2.} First, we will use the CSEW, and a logistic approach to understand how the decision to report to the police (or seeking any external help) depends on the individual-level characteristics of the victim (especially focusing on mental and physical health, financial independence and whether the victim has children) and the abuse suffered. We can complement this analysis with the WMP data,
